Storm surge modelling in the Strait of Georgia

### 2014-10-14 RoughWork/manuscript.tex
Included statistics for Dec 2006

The maximum water level and maximum residuals are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures in Table \ref{tab:feb2006stat}. In general, $\gamma^2$ and $WS$ suggest good model agreement for the predicted water levels, most likely because the tidal signal dominates. The skill scores degrade slightly for the residuals but are still reasonable and indicate very good agreement. Note that the mean absolute error ($\bar{e}$) is lower for the residuals, but this measure does not account for the larger variablity in water levels as opposed to residuals.

\begin{table}[h]
\centering 
\begin{tabular}{|l |c c c | c c c|} 
\hline 
& \multicolumn{3}{|c|}{Water Level ($\eta$)}        & \multicolumn{3}{|c|}{Residual ($\delta$)} \\ 
\hline 
Location       & $\bar{e}$ (cm) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\gamma^2$ & $WS$      \\
\hline 
Point Atkinson & 10.5           &   0.012    & 0.994  & 6.2            &  0.0646    & 0.979     \\
Victoria       &  8.6           &   0.032    & 0.992  & 6.1            &  0.0851    & 0.970     \\
Patricia Bay   &  9.1           &   0.024    & 0.992  & 6.9            &  0.0704    & 0.969     \\
Campbell River & 13.1           &   0.041    & 0.990  & 5.8            &  0.0790    & 0.961     \\
\hline 
\end{tabular}
\caption{Evaluation of model performance for the Feb 2006 simulation. Mean absoulte error ($\bar{e}$), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and resdiuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006}.}
\label{tab:feb2006stat} 
\end{table} 
%anything else to include? statistics calculated in Feb 2006 -paper prodcution.ipynb and saved in statisitcs_feb2006.csv

@@

\subsection{Spatial Extent of Surge: Dec 15, 2006}
On Dec 15, 2006, a major storm with wind gusts over 120 km/hr caused significant damage to many coastal communities on the Pacific Northwest. Although the storm was very strong and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area\citep{forseth2006adaptation}. This event has been used to examine the propagation of the surge throughout the domain. First, an assessment of the accuracy of the hindcast is provided.

Point Atkinson observed a maximum surge amplitude of 84 cm, which is approximately 29 cm higher than the model result, however the timing of the surge agrees almost exactly. as displayed in Figure \ref{fig:dec2006} The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 44 cm. At all of the locations, the model predicted the maximum surge to occur within 3.5 hours of the observed surge. Typically, the modelled surge was too early, in contrast to the later predictions in the Feb 4, 2006 case. Errors in the modelled water level at the time of maximum surge were of the same order as the errors in the surge amplitude.

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/dec2006.pdf}
\caption{Comparison of observed water level and surge amplitudes during the Dec 15, 2006 storm surge simulation. }
\label{fig:dec2006}
\end{figure}%table instead of figure?

>>>>

The maximum water level and maximum residuals are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures in Table \ref{tab:feb2006stat}. In general, $\gamma^2$ and $WS$ suggest good model agreement for the predicted water levels, most likely because the tidal signal dominates. The skill scores degrade slightly for the residuals but are still reasonable and indicate very good agreement. Note that the mean absolute error ($\bar{e}$) is lower for the residuals, but this measure does not account for the larger variablity in water levels due to the tides.

\begin{table}[h]
\centering 
\begin{tabular}{|l |c c c | c c c|} 
\hline 
& \multicolumn{3}{|c|}{Water Level ($\eta$)}        & \multicolumn{3}{|c|}{Residual ($\delta$)} \\ 
\hline 
Location       & $\bar{e}$ (cm) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\gamma^2$ & $WS$      \\
\hline 
Point Atkinson & 10.5           &   0.012    & 0.994  & 6.2            &  0.065     & 0.979     \\
Victoria       &  8.6           &   0.032    & 0.992  & 6.1            &  0.085     & 0.970     \\
Patricia Bay   &  9.1           &   0.024    & 0.992  & 6.9            &  0.070     & 0.969     \\
Campbell River & 13.1           &   0.041    & 0.990  & 5.8            &  0.079     & 0.961     \\
\hline 
\end{tabular}
\caption{Evaluation of model performance for the Feb 2006 simulation. Mean absoulte error ($\bar{e}$), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and resdiuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006}.}
\label{tab:feb2006stat} 
\end{table} 
%anything else to include? statistics calculated in Feb 2006 -paper prodcution.ipynb and saved in statisitcs_feb2006.csv 
%mean of model and mean of observations?

@@

\subsection{Spatial Extent of Surge: Dec 15, 2006}
On Dec 15, 2006, a major storm with wind gusts over 120 km/hr caused significant damage to many coastal communities on the Pacific Northwest. Although the storm was very strong and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}. This event has been used to examine the propagation of the surge throughout the domain and spatial variations in the surge amplitude. First, an assessment of the accuracy of the hindcast is provided.

Point Atkinson observed a maximum surge amplitude of 84 cm, which is approximately 28 cm higher than the modelled residual, however the timing of the surge agrees almost exactly as displayed in Figure \ref{fig:dec2006}. The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 44 cm. At all of the locations, the model predicted the maximum surge to occur within 3.5 hours of the observed surge. Typically, the modelled surge was too early, in contrast to the later predictions in the Feb 4, 2006 case. Error and skill scores for the modelled water level and the modelled residual are provided in Table \ref{dec2006stat}. The modelled residuals are on average too low and have a lower skill score than the Feb 2006 case. This could be due to a number of factors. First,the very large wind speeds observed during this storm are significantly underrepresented by the atmospheric model leading to lower wind stress and less piling of water against the coast. Second, the anomaly forcing at the open boundaries could be too low. The observed residual at Tofino, BC (a point near by but outside of the domain) is up to 10 cm lower than the observed residual at Neah Bay, WA (a point near by but inside of the domain). Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the case. 
%Skill score between CGRF and observed winds?

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/dec2006.pdf}
\caption{Comparison of observed and modelled water level and residuls during the Dec 15, 2006 storm surge simulation. }
\label{fig:dec2006}
\end{figure}%table instead of figure?

\begin{table}[h]
\centering 
\begin{tabular}{|l |c c c | c c c|} 
\hline 
& \multicolumn{3}{|c|}{Water Level ($\eta$)}        & \multicolumn{3}{|c|}{Residual ($\delta$)} \\ 
\hline 
Location       & $\bar{e}$ (cm) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\gamma^2$ & $WS$      \\
\hline 
Point Atkinson & 10.5           &   0.025    & 0.994  & 11.1           &  0.275     & 0.891     \\
Victoria       &  9.6           &   0.042    & 0.984  &  7.9           &  0.223     & 0.911     \\
Patricia Bay   &  9.2           &   0.031    & 0.990  &  9.5           &  0.260     & 0.889     \\
Campbell River & 10.6           &   0.030    & 0.992  & 10.6           &  0.247     & 0.905     \\
\hline 
\end{tabular}
\caption{Evaluation of model performance for the Dec 2006 simulation.}
\label{tab:dec2006stat} 
\end{table}
